---
title: My dynamic report
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)

```

```{css}
@import url('https://fonts.googleapis.com/css2?family=Quicksand:wght@300..700&family=Shrikhand&display=swap');

h1{
font-family: "Shrikhand", serif;
font-size: 2.5em;
}
img{
    border-style: solid;
    border-width: 5px;
    border-color: #F2B5D4;
    border-radius: 20px;
}
h2{
font-family: "Shrikhand", serif;

}
p{
    font-family: "Quicksand", sans-serif;
}
body {
  background-image: linear-gradient(#FFFFFF,#EFF7F6, #30BCED);
} 
```

# Auckland Transport Punctuality

I normally take public transport into University, I would like to know on average how delayed the transport is and if different modes of public transport have more delay compared to others. 

Note: I did not take all of the trips, some were taken by friends and some I just logged from looking at the signs at the stops.

```{r}
library(tidyverse)
# reading in the csv into logged data

csv_file <- "https://docs.google.com/spreadsheets/d/e/2PACX-1vThmBZ64dA12eqXns5uIpIwGp0geQ3BTaVLhnifiGhNVvAUQj6ev5uUhdsDh8FdAFLt1mrxYrsW33oY/pub?output=csv"
logged_data <- read_csv(csv_file)

# Renaming the variables in the data frame latest_data

latest_data <- logged_data %>%
  rename(transport_type = 2,
         transport_time = 3,
         transport_delay = 4)
# Summary Values

# Finding the mean of transport_delay
mean_delay <- round(mean(latest_data$transport_delay), 2)

# Finding the maximum amount of delay

max_delay <- max(latest_data$transport_delay)

# A Bar Chart showing the different types of public transport taken
transport_taken <-
ggplot(data = latest_data) +
  geom_bar(aes(x = transport_type,
           fill = transport_type)) +
  labs(title = "What Kind of Public Transport are people taking?", 
       subtitle = "The different kinds of public transport are using in auckland from my survey",
       y = "Trips Taken",
       x = "Types of Transport") +
  scale_fill_manual(values = c("#F2B5D4","#AF125A", "#30BCED"))

# A Bar Chart of the delay in minutes, by what transport was taken
bar_chart1 <-
ggplot(data = latest_data) + 
  geom_bar(aes(x = transport_delay, fill = transport_type)) +
  labs(title = "Are different types of public Transport more Reliable?",
       subtitle = "Comparing the delay and the types of transport in Auckland",
       x = "Delay in minutes",
       y = "Number of Trips") +
    scale_fill_manual(values = c("#F2B5D4","#AF125A", "#30BCED"))


```

## Overall Delay
Overall the average amount of delay in minutes for the trips I recorded was `r mean_delay`, The max amount of delay is `r max_delay` minutes. In a bar chart it looks like this:

```{r}
overall_lateness <- 
  ggplot(data = latest_data) + 
  geom_bar(aes(x = transport_delay), fill = "#F2B5D4") + 
  labs(title = "Delay of all transport taken", 
       x = "Delay in minutes", 
       y = "Frequency")

overall_lateness

bus_df <- latest_data %>% filter(transport_type == "Bus")
bus_mean <- round(mean(bus_df$transport_delay),2)

# finding the mean of the trains 
train_df <- latest_data %>% filter(transport_type == "Train")
train_mean <- round(mean(train_df$transport_delay), 2)

```

## Delay by type of Transport

I also wanted to know what type of public transport is the most reliable. The mean for buses was `r bus_mean` minutes compared to the mean of trains being `r train_mean` minutes late. An easier way to see this is through the bar chart below

```{r}
ggplot(data = latest_data) + 
  geom_bar(aes(x = transport_delay, fill = transport_type)) +
  labs(title = "Are different types of public Transport more Reliable?",
       subtitle = "Comparing the delay and the types of transport in Auckland",
       x = "Delay in minutes",
       y = "Number of Trips") +
  scale_fill_manual(values = c("#F2B5D4","#AF125A", "#30BCED"))

```

However this data may not be entirely accurate as there was a difference in how many trips taken I recorded on each Transport type, This most likely skews the data. 

```{r}
transport_taken

morning_mean <- latest_data %>% 
  filter(transport_time == "1 - 10") %>%
  pull(transport_delay) %>%
  mean %>%
  round(2)

midday_mean <- latest_data %>%
  filter(transport_time == "10 - 4") %>%
  pull(transport_delay) %>% 
  mean %>%
  round(2)

evening_mean <- latest_data %>%
  filter(transport_time == "4 - 10") %>%
  pull(transport_delay) %>%
  mean %>%
  round(2)


```

## Does the time of day effect how delayed the Transport is?
Another thing I wanted to see was if depending on the time of day would there be more or less delay.

```{r}
paste0("The mean delay for morning (1am to 10am) is ", morning_mean, " The mean delay for midday(10 am to 4pm) is ", midday_mean, " and the evening(4pm to 10pm) mean is ", evening_mean, ".")

lateness_time <-
ggplot(data = latest_data) +
  geom_bar(aes(x = transport_delay, fill = transport_time)) +
  labs(title = "Does the time of day effect delay?",
       subtitle = "Amount of Delay in minutes by what time of day the trip was taken", 
       x = "Amount of delay in minutes",
       y = "Trips Taken") +
  scale_fill_manual(values = c("#F2B5D4","#AF125A", "#30BCED"))


lateness_time
```


